---
title: "P8106_hw2"
author: "Nathalie Fadel"
date: "3/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Import data
```{r}
library(tidyverse)
concrete = read_csv("Data/concrete.csv") %>%
  janitor::clean_names()

```

###(a) Create scatter plots of response vs. predictors using the function featurePlot().
```{r, fig.height=5}
library(caret)

# matrix of predictors 
x <- model.matrix(compressive_strength~.,concrete)[,-1]
# vector of response
y <- concrete$compressive_strength

theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","compressive strength"),
            type = c("p"), layout = c(4, 2))
```

###(b) Perform polynomial regression to predict compressive strength using water as the predictor. For 1 ≤ d ≤ 4, use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the results of hypothesis testing using ANOVA? Make a plot of different polynomial fits to the data.
```{r}
library(glmnet)
library(ISLR)
library(mgcv)
library(boot)

fit1 <- lm(compressive_strength~water, data = concrete) #1st degree
fit2 <- lm(compressive_strength~poly(water,2), data = concrete) #squared
fit3 <- lm(compressive_strength~poly(water,3), data = concrete) #cubed
fit4 <- lm(compressive_strength~poly(water,4), data = concrete) #4th power

anova(fit1,fit2,fit3,fit4) 

#finding optimal exponent  
set.seed(1)
mseK <- rep(NA, 4)
for (i in 1:4) {
    fit <- glm(compressive_strength ~ poly(water, i), data = concrete)
    mseK[i] <- cv.glm(concrete, fit, K = 10)$delta[1]
}
plot(1:4, mseK, xlab = "Exponent", ylab = "Test MSE", type = "l")


#scatter plot with fit lines  
plot(compressive_strength ~ water, data = concrete, col = "blue")

waterlims <- range(concrete$water)
water.grid <- seq(from = waterlims[1], to = waterlims[2], by = 1)

preds1 <- predict(fit1, newdata = data.frame(water = water.grid))
preds2 <- predict(fit2, newdata = data.frame(water = water.grid))
preds3 <- predict(fit3, newdata = data.frame(water = water.grid))
preds4 <- predict(fit4, newdata = data.frame(water = water.grid))

lines(water.grid, preds1, col = "red", lwd = 2)
lines(water.grid, preds2, col = "purple", lwd = 2)
lines(water.grid, preds3, col = "green", lwd = 2)
lines(water.grid, preds4, col = "brown", lwd = 2)
```  
A degree 4 polynomial is the best fit for the data, according to both the ANOVA test and the 10-fold cross-validation.

###(c) Fit a smoothing spline using water as the predictor for a range of degrees of freedom, as well as the degree of freedom obtained by generalized cross-validation, and plot the resulting fits. Describe the results obtained.
```{r}
library(splines)

#Generating Test Data
waterlims <- range(concrete$water)
water.grid <- seq(from=waterlims[1], to = waterlims[2])

#fit 1
fit.ss1 <- smooth.spline(concrete$water, concrete$compressive_strength, df = 10)
fit.ss1
pred.ss1 <- predict(fit.ss1, x = water.grid)
pred.ss.df1 <- data.frame(pred = pred.ss1$y, water = water.grid)

#fit 2
fit.ss2 <- smooth.spline(concrete$water, concrete$compressive_strength, df = 30)
fit.ss2
pred.ss2 <- predict(fit.ss2, x = water.grid)
pred.ss.df2 <- data.frame(pred = pred.ss2$y, water = water.grid)

#fit 3
fit.ss3 <- smooth.spline(concrete$water, concrete$compressive_strength, df = 50)
fit.ss3
pred.ss3 <- predict(fit.ss3, x = water.grid)
pred.ss.df3 <- data.frame(pred = pred.ss3$y, water = water.grid)

#cv fit
fit.ss <- smooth.spline(concrete$water, concrete$compressive_strength)
fit.ss
pred.ss <- predict(fit.ss, x = water.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, water = water.grid)


#plot of fits
scatter <- ggplot(data = concrete, aes(x = water, y = compressive_strength)) +
  geom_point(color = rgb(.2, .4, .2, .5))

smooth_fits <- scatter + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df1,
          color = rgb(.8, .1, .1, 1)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df2,
          color = rgb(0, 0, 1, 1)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df3,
          color = rgb(1, 0, 1, 1)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df,
          color = rgb(.2, .2, .4, 1)) + 
  theme_bw()

smooth_fits
```
The degrees of freedom obtained by cross-validation is approximately 69. The best fit line using this df value has many local maxima and minima, and is subject to more local variation. With a low df, the fit line is smoother, but it may not capture local variation as well.  

###(d) Fit a GAM using all the predictors. Plot the results and explain your findings.
```{r}
gam.m1 <- gam(compressive_strength ~ cement + blast_furnace_slag + fly_ash + water + superplasticizer + coarse_aggregate + fine_aggregate + age, data = concrete)
gam.m2 <- gam(compressive_strength ~ cement + blast_furnace_slag + fly_ash + s(water) + superplasticizer + coarse_aggregate + fine_aggregate + s(age), data = concrete)

plot(gam.m2)

anova(gam.m1, gam.m2, test = "F")
```  
In the model, we have added age as a smoothing term, in addition to water. For the graphs, age shows very low standard error around 0 but higher SE as age increases. For water, there is fairly low SE throughout, and is fairly limited to the ends of the graph. The ANOVA test shows that the model with age and water as smoothing terms is significantly different than the simple GAM with no smoothing. 